Astro 507; Problem Set 3

### 2022-02-25 pset3/solutions.tex
2b

\question{1. \textbf{FIRAS}}

\answer{}

@@

\question{2b. \textbf{Expanding pressure}}
\answer{
    \begin{align}
        \frac{P}{n k_B T} &= \frac{4(2s + 1)}{3 \pi^{1/2}} k_B T \lambda^{-3} F_{3/2}(z) \qty[\frac{2(2s + 1)}{\pi^{1/2} \lambda^3} k_B T F_{1/2}(z)]^{-1} \\
                          &= \frac{2 F_{3/2}(z)}{3 F_{1/2}(z)}
    \end{align}
    \todo{come back to this on Thursday}
}

>>>>

\question{1. \textbf{FIRAS}}
\answer{
    \todo{copy over stuff from jupyter notebook}
}

@@

\question{2b. \textbf{Expanding pressure}}
\answer{
    To start, we can right a simplified expression for $P / n k_B T$ using the expressions from part a.
    \begin{align}
        \frac{P}{n k_B T} &= \frac{4(2s + 1)}{3 \pi^{1/2}} k_B T \lambda^{-3} F_{3/2}(z) \qty[\frac{2(2s + 1)}{\pi^{1/2} \lambda^3} k_B T F_{1/2}(z)]^{-1} \\
        \frac{P}{n k_B T} &= \frac{2 F_{3/2}(z)}{3 F_{1/2}(z)}
    \end{align}
    Now we need to expand this in terms of $z$. Let's start by just expanding the Fermi-Dirac integral to start with. We know this is defined as
    \begin{equation}
        F_v(z) = \int_0^\infty \dd{w} \frac{w^v}{e^w z^{-1} + 1}
    \end{equation}
    If we perform a Taylor expansion on this we find
    \begin{equation}
        F_v(z) \approx \int_0^\infty \qty(\sum_{i = 1}^{\infty} e^{- i w} w^v z^i) \dd{w}
    \end{equation}
    We can switch the order of this sum and integral and also pull out the $z$ terms to find that
    \begin{equation}
        F_v(z) \approx \sum_{i = 1}^{\infty} \qty( z_i \int_0^\infty e^{- i w} w^v \dd{w} )
    \end{equation}
    Let's keep the first two terms of this (since just keeping the first results in just getting 1 as the answer haha). This gives the ratio as
    \begin{equation}
        \frac{P}{n k_B T} \approx \frac{2}{3} \frac{\frac{3 \sqrt{\pi}}{4} z + \frac{3 \sqrt{\pi}}{16 \sqrt{2}} z^2}{\frac{\sqrt{\pi}}{2} z + \frac{\sqrt{\pi}}{4 \sqrt{2}} z^2}
    \end{equation}
    Also of those factors of $\sqrt{\pi}$ cancel and we can also taylor expand again to get a nice simple expression as a function of the fugacity
    \begin{equation}
        \boxed{ \frac{P}{n k_B T} \approx 1 + \frac{z}{4 \sqrt{2}} }
    \end{equation}
}
